Tweet Classification: Filtering \\ of Twitter for Company-relevant Tweets

### 2012-12-07 paper/paper.tex
added failing paper
paper
paper updates

\begin{table}[h]
\centering
\begin{tabular}{|l|}
	\hline
\textbf{@Apple} please make a squid \textbf{\#emoji} so I can put \\
one next to \textbf{@squidneyy22}'s name\\ \hline
Maybe \textbf{\#Apple} will release \textbf{\#iTunes11} \\
tomorrow. Maybe? \\ \hline
\end{tabular}
\caption{Hashtags and username references (note that Apple Inc. does not own the handle @Apple, though it is commonly used this way)}
\label{tab:myfirsttable}
\end{table}

@@

\begin{table}[h]
\centering
\begin{tabular}{|l|}
	\hline
\textbf{Nvm} i'll buy you an ipad \textbf{\^-\^} \\ \hline
\#Apple fires `maps' project manager \textbf{loooool} big \\
 news story here in San Francisco! Go figure. \textbf{:-)} \\ \hline
Think I might have a \textbf{gf}! \textbf{O\_o} oh \\
lord\textbf{..} Lets see how this goes. \\ \hline
\end{tabular}
\caption{Examples of emoticons and common abbreviations found on Twitter}
\label{tab:myfirsttable}
\end{table}

@@

\subsubsection{Feature-space Enrichment} ~\\
\textit{Stopword removal}: We used Mallet's TokenSequenceRemoveStopwords Pipe to remove 500 of the most common words in the English language from each tweet. We remove these words because of the high likelihood of many tweets containing these words, without contributing significant discriminatory features to the tweets. %TODO [REFERENCE NLP PAPER ABOUT STOPWORD REMOVAL]
\\
~\\
\textit{Bigrams}: Using n-grams is a common method of increasing the number of features present in short text%[FIND CITATION]
. Often, this is enough to increase accuracy significantly%CITATION
. As discussed in the tokenization section below, we used selective bigramming, which yielded increases in accuracy in almost all test cases.\\
~\\
\textit{HTML corrections}: When accessing tweets through Twitter's public API, the native JSON format includes HTML-encoded entities that can add confusing features to the tweets if not un-escaped. For instance, angle brackets are replaced with \&lt; and \&gt; (less-than and greater-than), causing emoticons, for instance, to no longer be parseable. The emoticon `\textgreater.\textless', for instance, would be `\&gt;.\&lt;'. Necessarily, this Pipe un-escapes all such entities, making them easier to parse.\\
~\\
\textit{URL replacement}: In our dataset, 60\% of the 100,000 tweets we collected contained a link. As the vast majority of these are unique, and shortened using a URL shortener, often Twitter's own t.co. Links, then, lose any features that may have been parsed from the URL, which often contains a domain name or title that may be useful in discriminating especially ambiguous links. Consider, for instance, the tweet `hmmm http://t.co/39D8I3qM' which leaves a parser completely unaware of the true content of the tweet. Following this URL leads to `http://www.redmondpie.com/iphone-5-vs-iphone-4s-should-you-upgrade/', which is clearly related to Apple. This pipe replaces the link itself in the tweet with the title of the destination web page: `iPhone 5 vs iPhone 4S: Should You Upgrade? | Redmond Pie.' In cases where the title is empty, we instead replace the shortened link with the resolved URL, which is `http://www.redmondpie.com/iphone-5-vs-iphone-4s-should-you-upgrade' in this case. In many cases like this, there is not other way to distinguish the content of a tweet. From our dataset of 100,000 tweets, 102 tweets contained \textit{only} a link, while 608 tweets contained only a link and fewer than 20 other characters. While this is a fairly small portion of the overall dataset for training purposes, it can be crucial in making decision on novel test instances. Pak and Paroubek also noted the need to handle URLs in tweets, however, rather than adding any new text, they chose to remove the link entirely  (\citeauthor{PAK10.385}).
~\\
\textit{Spell checking}: Twitter's user-generated content contains a large volume of misspelled words (\citeauthor{Laboreiro:2010:TMM:1871840.1871853}). Using a spell-checking library, we wrote a Pipe to replace unrecognized words (misspelled words) with their corrections if the confidence for correction was above a certain threshold. Our intention was to reduce the noise in the feature set by eliminating common misspellings.
~\\
\subsubsection{Disambiguation}
~\\
\textit{Stemming}: Stemming is the process of reducing words to their `stem,' in order to group different forms of the same word into a single feature. While this stem does not necessarily have to be the root of the word, it should satisfy the property that related words have the same stem. For instance, the stem of the related words `argue,' `argued,' and `arguing' all have the stem of `argu,' even though argu is not the true linguistic stem, or a real word. We used the standard Porter stemmer to implement this (\citeauthor{porter_1980}).
In some cases, stemming proved to increase our accuracy.
\\

\begin{figure*}[t!]
\centering
\includegraphics[height=7cm]{img/all-with-headings}
\caption{Metrics for all Pipeline combinations, training set sizes 50 through 1500 instances}
\label{fig:chart_all}
\end{figure*}

@@

\section{Experimental Setup}
In order to determine which combination of preprocessing steps would result in the highest performance, we built a custom evaluation utility that iterates through all possible valid combinations of pipes and evaluates their performance. This led to nearly one hundred combinations, as each of six tokenizer variations was run with and without bigramming, with and without stemming, etc. For consistency, the disjoint training and test sets used to evaluate each combination were identical for each iteration. To best observe the effects of training set size, training sizes of 50 to 1500 were tested in increments of 50, where each set is a superset of the previous set with 50 new randomly chosen instances, as opposed to using n-fold cross validation, which would not display this characteristic evenly across varied training and testing sizes. To evaluate this large set of results, we have chosen three relevant evaluation metrics. The first is precision, or the ratio of correctly labeled tweet instances to the number of instances assigned a particular label (`apple' or `none' in our case). Higher values for this metric mean fewer false positives for the desired company, which is the overall goal in this classification task. The second metric was recall, or the ratio of correctly classified instances out of the full set of instances which should have the label in question. This metric is less relevant to our purpose, but still interesting. The sheer scale of Twitter means missing even half of the tweets mentioning a company and mislabeling them as about `none' is of little consequence. Finally we use the F1 metric, a standard combination of precision and recall that is useful in determining overall if some pipeline combination is better than another in both precision and recall.

@@

Our desired result was a high precision classifier that could be useful gathering a large number of tweets mentioning a company or its products. Looking at an overview of the results in Figure~\ref{fig:chart_all}, we can see all Pipeline combinations had over 70\% of ``apple'' labelings correct with only 50 training tweet instances. However, given that a baseline strategy of labeling every tweet as ``apple'' would be correct 64\% of the time, this is only an improvement of approximately 9\% in the worst case. The more relevant and interesting result is the 43\% improvement from the baseline seen at

\begin{table}[h]
\centering
\begin{tabular}{|l|r|}
	\hline
	Pipeline Description & Apple Precision \\ \hline
	Twitter Tokenizer, Lowercased, Stemmed, \\ Stopwords Removed, Bigrams & 0.9224489796 \\ \hline
	% TODO FIIIIIIIX
	[Input2CharSequence, UnescapeHTML, CharSequenceLowercase, default twitter tokenizer, TokenSequenceRemoveStopwords, NGRAM2, TokenSequence2FeatureSequence, Target2Label, FeatureSequence2FeatureVector] & 0.9159663866 \\ \hline
	[Input2CharSequence, UnescapeHTML, default twitter tokenizer, TokenSequenceRemoveStopwords, Stemmed, NGRAM2, TokenSequence2FeatureSequence, Target2Label, FeatureSequence2FeatureVector] & 0.9156626506 \\ \hline
	[Input2CharSequence, UnescapeHTML, default twitter tokenizer, Stemmed, NGRAM2, TokenSequence2FeatureSequence, Target2Label, FeatureSequence2FeatureVector] & 0.9138576779 \\ \hline
	[Input2CharSequence, UnescapeHTML, default twitter tokenizer, TokenSequenceRemoveStopwords, NGRAM2, TokenSequence2FeatureSequence, Target2Label, FeatureSequence2FeatureVector] & 0.9135802469 \\
	\hline
\end{tabular}
\caption{Top performing Pipeline combinations for 1500 training instances}
\label{tab:top_pipelines}
\end{table}

@@

\subsection{Pipe Breakdown}

\textit{Stopword removal}:
\textit{Bigrams}:
\textit{URL Replacement}:
~\\
\textit{Spell checking}: Correcting misspelled words provided no statistically significant change in our results. \\

@@

>>>>

\begin{table}[h]
\centering
\begin{tabular}{|l|}
	\hline
\textbf{@Apple} please make a squid \textbf{\#emoji} so I can put \\
one next to \textbf{@squidneyy22}'s name\\ \hline
Maybe \textbf{\#Apple} will release \textbf{\#iTunes11} \\
tomorrow. Maybe? \\ \hline
\end{tabular}
\caption{Hashtags and username references (note that Apple Inc. does not own the handle @Apple, though it is commonly used this way)}
\label{tab:hashtag_username}
\end{table}

@@

\begin{table}[h]
\centering
\begin{tabular}{|l|}
	\hline
\textbf{Nvm} i'll buy you an ipad \textbf{\^-\^} \\ \hline
\#Apple fires `maps' project manager \textbf{loooool} big \\
 news story here in San Francisco! Go figure. \textbf{:-)} \\ \hline
Think I might have a \textbf{gf}! \textbf{O\_o} oh \\
lord\textbf{..} Lets see how this goes. \\ \hline
\end{tabular}
\caption{Examples of emoticons and common abbreviations found on Twitter}
\label{tab:emoticons}
\end{table}

@@

\subsubsection{Feature-space Enrichment} ~\\
\textit{Stopword removal}: We used Mallet's TokenSequenceRemoveStopwords Pipe to remove 500 of the most common words in the English language from each tweet. We remove these words because of the high likelihood of many tweets containing these words, without contributing significant discriminatory features to the tweets. %TODO [REFERENCE NLP PAPER ABOUT STOPWORD REMOVAL]
\\
~\\
\textit{Bigrams}: Using n-grams is a common method of increasing the number of features present in short text%[FIND CITATION]
. Often, this is enough to increase accuracy significantly%CITATION
. As discussed in the tokenization section below, we used selective bigramming, which yielded increases in accuracy in almost all test cases.\\
~\\
\textit{HTML corrections}: When accessing tweets through Twitter's public API, the native JSON format includes HTML-encoded entities that can add confusing features to the tweets if not un-escaped. For instance, angle brackets are replaced with \&lt; and \&gt; (less-than and greater-than), causing emoticons, for instance, to no longer be parseable. The emoticon `\textgreater.\textless', for instance, would be `\&gt;.\&lt;'. Necessarily, this Pipe un-escapes all such entities, making them easier to parse.\\
~\\
\textit{URL replacement}: In our dataset, 60\% of the 100,000 tweets we collected contained a link. As the vast majority of these are unique, and shortened using a URL shortener, often Twitter's own t.co. Links, then, lose any features that may have been parsed from the URL, which often contains a domain name or title that may be useful in discriminating especially ambiguous links. Consider, for instance, the tweet `hmmm http://t.co/39D8I3qM' which leaves a parser completely unaware of the true content of the tweet. Following this URL leads to `http://www.redmondpie.com/iphone-5-vs-iphone-4s-should-you-upgrade/', which is clearly related to Apple. This pipe replaces the link itself in the tweet with the title of the destination web page: `iPhone 5 vs iPhone 4S: Should You Upgrade? | Redmond Pie.' In cases where the title is empty, we instead replace the shortened link with the resolved URL, which is `http://www.redmondpie.com/iphone-5-vs-iphone-4s-should-you-upgrade' in this case. In many cases like this, there is not other way to distinguish the content of a tweet. From our dataset of 100,000 tweets, 102 tweets contained \textit{only} a link, while 608 tweets contained only a link and fewer than 20 other characters. While this is a fairly small portion of the overall dataset for training purposes, it can be crucial in making decision on novel test instances. Pak and Paroubek also noted the need to handle URLs in tweets, however, rather than adding any new text, they chose to remove the link entirely  (\citeauthor{PAK10.385}). \\
~\\
\textit{Spell checking}: Twitter's user-generated content contains a large volume of misspelled words (\citeauthor{Laboreiro:2010:TMM:1871840.1871853}). Using a spell-checking library, we wrote a Pipe to replace unrecognized words (misspelled words) with their corrections if the confidence for correction was above a certain threshold. Our intention was to reduce the noise in the feature set by eliminating common misspellings.
~\\
\subsubsection{Disambiguation}
~\\
\textit{Stemming}: Stemming is the process of reducing words to their `stem,' in order to group different forms of the same word into a single feature. While this stem does not necessarily have to be the root of the word, it should satisfy the property that related words have the same stem. For instance, the stem of the related words `argue,' `argued,' and `arguing' all have the stem of `argu,' even though argu is not the true linguistic stem, or a real word. We used the standard Porter stemmer to implement this (\citeauthor{porter_1980}).
In some cases, stemming proved to increase our accuracy.
\\

\begin{figure*}[t!]
\centering
\includegraphics[width=\textwidth]{img/all-with-headings}
\caption{Metrics for all Pipeline combinations, training set sizes 50 through 1500 instances}
\label{fig:chart_all}
\end{figure*}

@@

\section{Experimental Setup}
In order to determine which combination of preprocessing steps would result in the highest performance, we built a custom evaluation utility that iterates through all possible valid combinations of pipes and evaluates their performance. This led to nearly one hundred combinations, as each of six tokenizer variations was run with and without bigramming, with and without stemming, etc. For consistency, the disjoint training and test sets used to evaluate each combination were identical for each iteration. To best observe the effects of training set size, training sizes of 50 to 1500 were tested in increments of 50, where each set is a superset of the previous set with 50 new randomly chosen instances, as opposed to using n-fold cross validation, which would not display this characteristic evenly across varied training and testing sizes. \\

To evaluate this large set of results, we have chosen three relevant evaluation metrics. The first is precision, or the ratio of correctly labeled tweet instances to the number of instances assigned a particular label (`apple' or `none' in our case). Higher values for this metric mean fewer false positives for the desired company, which is the overall goal in this classification task. The second metric was recall, or the ratio of correctly classified instances out of the full set of instances which should have the label in question. This metric is less relevant to our purpose, but still interesting. The sheer scale of Twitter means missing even half of the tweets mentioning a company and mislabeling them as about `none' is of little consequence. Finally we use the F1 metric, a standard combination of precision and recall that is useful in determining overall if some pipeline combination is better than another in both precision and recall.

@@

Our desired result was a high precision classifier that could be useful gathering a large number of tweets mentioning a company or its products. Looking at an overview of the results in Figure~\ref{fig:chart_all}, we can see all Pipeline combinations had over 70\% of ``apple'' labelings correct with only 50 training tweet instances. However, given that a baseline strategy of labeling every tweet as ``apple'' would be correct 64\% of the time, this is only an improvement of approximately 9\% in the worst case. The more relevant and interesting result is the 43\% improvement from the baseline seen with our custom Twitter Tokenizer and all other Pipes like Lowsercasing, Stemming and Bigrams trained on the maximum of 1500 instances. This relatively large improvement indicates not only is the tokenization strategy used important, but also that we were relatively successful in creating an effective tokenizer that can be further augmented by standard natural language processing techniques like stemming, removing stop words, applying n-grams, or simply lowercasing the text. See in Table \ref{tab:top_pipelines} below that all of top five Pipeline combinations include our twitter tokenizer. In fact, one must go down to the 24th and 33rd ranked combinations to find ones that include the Number-Digit and Whitespace based tokenizers respectively.

\begin{table}[h]
\centering
\small
\begin{tabular}{|c|l|r|}
	\hline
	\textit{Rank} & \textit{Pipeline Description} & \textit{Precision} \\
	& & \textit{for Apple} \\ \hline
	1 & Twitter Tokenizer, Lowercased, & \\ & Stemmed, Stopwords Removed, Bigrams & 92.2\% \\ \hline
	% TODO FIIIIIIIX
	2 & Twitter Tokenizer, Lowercased, & \\ & Stopwords Removed, Bigrams & 91.6\% \\ \hline
	3 & Twitter Tokenizer, Stopwords Removed, & \\ & Stemmed, Bigrams & 91.6\% \\ \hline
	4 & Twitter Tokenizer, Stemmed, Bigrams & 91.4\% \\ \hline
	5 & Twitter Tokenizer, Stopwords Removed, & \\ & Bigrams & 91.3\% \\ \hline
	\hline
	24 & Number-Digit Tokenizer, & \\ & Remove Stopwords, Bigrams & 88.2\%
 \\ \hline
	\hline
	33 & Whitespace Tokenizer, Stemmed, & \\ & Remove Stopwords, Bigrams & 87.7\% \\ \hline

\end{tabular}
\caption{Top performing Pipeline combinations for 1500 training instances}
\label{tab:top_pipelines}
\end{table}

@@

\subsection{Pipe Effectiveness Breakdown}

\textit{Stopword removal}: \\
~\\
\textit{Bigrams}: \\
~\\
\textit{URL Replacement}: \\
~\\
\textit{Spell checking}: Correcting misspelled words provided no statistically significant change in our results. \\

@@

\begin{table}[h]
    \begin{tabular}{|l|l|l|l|}
    \hline
        Apple  & i     & Apple     & apple        \\
        iphone & apple & iphone    & Apple        \\ 
        i      & the   & ...       & ...          \\ 
        the    & Apple & ipad      & mac          \\ 
        to     & and   & ipod      & juice        \\ 
        my     & to    & rt        & rt           \\ 
        a      & a     & \#apple    & post        \\ 
        ...    & ...   & 5         & new\_post    \\ 
        on     & for   & iphone\_5  & iphone      \\ 
        for    & new   & my\_iphone & apple\_juice\\
      \hline
    \end{tabular}
\end{table}
